---
title: "Region 5 State of the Child: Draft Data Findings and Visuals"
author: "Catalyst California"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    number_sections: no
    theme: sandstone
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=FALSE, message=FALSE, warning=FALSE)
library(usethis)

```

The following site holds the draft findings and draft visuals from various data sources to better understand the well-being and health of children in Best Start Region 5, the Antelope Valley. The findings and visuals are being developed by Catalyst California for the Region 5 State of the Child report in collaboration with The Children's Bureau. 

# Prep and Set Up 

```{r}
#### Loading Libraries ####
library(dplyr)
library(ggplot2)
library(RPostgreSQL)
library(tidyr)
library(data.table)
library(sysfonts)
library(showtext)

#### First 5 LA Style Guide ####
## COLORS## Taken from W:\Project\RDA Team\Region 5 State of the Child\Documentation\F5LA_BrandGuidelines_COLORS.pdf
#primary
lightblue <- "#009CDB"
darkblue <- "#332985"
tealblue <- "#22BCB8"
black <- "#000000"
textgrey <- "#919191"
#secondary 
green <- "#54B847"
orange <- "#F58326"
hotpink <- "#EC098C"
red <- "#EF4034"

## FONTS ##
# Step 1: Download fonts (Gotham Bold and Gotham Book) 
# Step 2: Load the downloaded fonts into the R script:
# General: See tutorial here, under "THE SHOWTEXT PACKAGE" section: https://r-coder.com/custom-fonts-r/#The_showtext_package 
# Step 3: Run the code below in each R script

font_add(family = "GothamBold", regular = "C:/Windows/Fonts/Gotham-Bold 700.otf")
font_add(family = "GothamBook", regular = "C:/Windows/Fonts/Gotham-Book 325.otf")

showtext_auto()

# define fonts in chart
font_title <- "GothamBold"
font_subtitle <- "GothamBook"
font_caption <- "GothamBook"
font_bar_label <- "GothamBold"
font_axis_label <- "GothamBook"
font_table_text<-"GothamBook"

#### Loading Database from PgAdmin/Postgres #### 
source("password_protection.R")

drv <- dbDriver("PostgreSQL")

con <- dbConnect(drv, dbname = "region5stateofthechild",
                   host = "aws-postgres-db.c5udgz7ro8hq.us-west-2.rds.amazonaws.com", port = 5432,
                   user = "postgres", password = pw)

#### Static Visual Functions ####

#Reference: W:\Project\RDA Team\LAFed\R\LAFed\visual_functions_static.R

```

# Introduction 

# Demographics 

## Population of Children and Families

The Antelope Valley is a 2,200 square mile area in northern Los Angeles County [1](https://en.wikipedia.org/wiki/Antelope_Valley). According to the US Census, it's two largest cities, Lancaster and Palmdale, have populations of 170,150 and 165,761, respectively.[1](https://www.census.gov/quickfacts/fact/table/palmdalecitycalifornia,lancastercitycalifornia/PST045221). For comparison, both cities are more populous than Pasadena and Pomona, two well-known mid-size cities on the other side of the valley. 

## Racial Ethnic Breakdown 

According to the latest American Community Survey data calculated by Lucy Wilkerson, Region 5 is majority Latinx. The regional share of the Black population is twice the LA County share and nearly three times the State share. On the other hand, the region is home to relatively fewer Asian residents than the County and State as a whole. 

```{r}
source("W:/Project/RDA Team/Region 5 State of the Child/R/race.R")

library(kableExtra)
kbl(race_data) %>%
  kable_styling()

```
Source: [American Community Survey 2016-2020 5-year estimates calculated for Antelope Valley neighborhoods by Lucy Wilkerson](https://public.tableau.com/app/profile/luz3725/viz/2020CensusData-AVBESTSTARTREGION5/GRAPI)

## Black Infant Mortality 

## Children of Incarcerated 

## Foster Youth Population 

## Student Homelessness

## Homelessness

The 2022 Los Angeles County homeless count counted over 65,000 homeless people in Los Angeles County, 1,000 of whom lived in the cities of Lancaster and Palmdale.

```{r}
source("W:/Project/RDA Team/Region 5 State of the Child/R/homeless.R")

library(kableExtra)
kbl(av_table) %>%
  kable_styling()

```
Source: [Los Angeles Homeless Services Authority, 2022](https://www.lahsa.org/data)


## Children with Special Needs 

## Languages of Children and Families

According to the latest American Community Survey data calculated by Lucy Wilkerson, the Antelope Valley has a diverse set of language speakers. The East Antelope Valley and Palmdale areas have a larger shares of Spanish speakers than the County or State as a whole. There are relatively fewer speakers of Asian or Pacific Islander languages, with West Antelope Valley having the largest share of this group. 
```{r}
source("W:/Project/RDA Team/Region 5 State of the Child/R/language.R")

library(kableExtra)
kbl(data) %>%
  kable_styling()

```
Source: [American Community Survey 2016-2020 5-year estimates calculated for Antelope Valley neighborhoods by Lucy Wilkerson](https://public.tableau.com/app/profile/luz3725/viz/2020CensusData-AVBESTSTARTREGION5/GRAPI)

# Resources(and barriers to access)

# Economics 

## Rent Burden (Cost Burden Renter)

## Housing Insecurity 

## Cost of Living

According to the United Ways of California, roughly half of households in and around Lancaster and Palmdale do not earn sufficient incomes to meet their basic costs of living (i.e., they earn incomes below the real cost measure of living). Lancaster and Palmdale adjusted median household incomes are well below other parts of the County and State.
```{r}
source("W:/Project/RDA Team/Region 5 State of the Child/R/cost_of_living.R")

library(kableExtra)
kbl(rcm_data) %>%
  kable_styling()

```
Source: [United Ways of California Real Cost Measure, 2021](https://www.unitedwaysca.org/realcost/39-real-cost)

## Food Insecurity 

Feeding America estimates that 22,740 children are food insecure[1](https://map.feedingamerica.org/). Over 10,000 Antelope Valley children under 5  participate in WIC. Latinx children comprise the largest share of Antelope Valley WIC participants, with nearly 1 in 5 Palmdale participants live in a Spanish-only home. Black children make up nearly 1 in 4 Lancaster WIC participants.
```{r}
source("W:/Project/RDA Team/Region 5 State of the Child/R/food_security.R")

library(kableExtra)
kbl(wic_data) %>%
  kable_styling()

```
Source: [L.A. County WIC Administrative Data, 2021](https://lawicdata.org/data-research/by-region/)

# Education

## Highschool Graduation

## Teacher & Staff Diversity 

## Chronic Absenteeism 

## Suspensions 

## Explusions 

## Early Childhood Edducation 

## 3rd Grade Math Scores

## 3rd Grade English Scores

# Safety and Education 

## Parks and Safe Spaces

## Police Violence

## Traffic Stops
