---
title: "Region 5 State of the Child: Draft Data Findings and Visuals"
author: "Catalyst California"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    number_sections: no
    theme: sandstone
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=FALSE, message=FALSE, warning=FALSE)
library(usethis)

```

The following site holds the draft findings and draft visuals from various data sources to better understand the well-being and health of children in Best Start Region 5, the Antelope Valley. The findings and visuals are being developed by Catalyst California for the Region 5 State of the Child report in collaboration with The Children's Bureau. 

# Prep and Set Up 

```{r}
#### Loading Libraries ####
library(dplyr)
library(ggplot2)
library(RPostgreSQL)
library(tidyr)
library(data.table)
library(sysfonts)
library(showtext)

#### First 5 LA Style Guide ####
## COLORS## Taken from W:\Project\RDA Team\Region 5 State of the Child\Documentation\F5LA_BrandGuidelines_COLORS.pdf
#primary
lightblue <- "#009CDB"
darkblue <- "#332985"
tealblue <- "#22BCB8"
black <- "#000000"
textgrey <- "#919191"
#secondary 
green <- "#54B847"
orange <- "#F58326"
hotpink <- "#EC098C"
red <- "#EF4034"

## FONTS ##
# Step 1: Download fonts (Gotham Bold and Gotham Book) 
# Step 2: Load the downloaded fonts into the R script:
# General: See tutorial here, under "THE SHOWTEXT PACKAGE" section: https://r-coder.com/custom-fonts-r/#The_showtext_package 
# Step 3: Run the code below in each R script

font_add(family = "GothamBold", regular = "C:/Windows/Fonts/Gotham-Bold 700.otf")
font_add(family = "GothamBook", regular = "C:/Windows/Fonts/Gotham-Book 325.otf")

showtext_auto()

# define fonts in chart
font_title <- "GothamBold"
font_subtitle <- "GothamBook"
font_caption <- "GothamBook"
font_bar_label <- "GothamBold"
font_axis_label <- "GothamBook"
font_table_text<-"GothamBook"

#### Loading Database from PgAdmin/Postgres #### 
source("password_protection.R")

drv <- dbDriver("PostgreSQL")

con <- dbConnect(drv, dbname = "region5stateofthechild",
                   host = "aws-postgres-db.c5udgz7ro8hq.us-west-2.rds.amazonaws.com", port = 5432,
                   user = "postgres", password = pw)

#### Static Visual Functions ####

#Reference: W:\Project\RDA Team\LAFed\R\LAFed\visual_functions_static.R

```

# Introduction 

# Demographics 

## Population of Children and Families

## Racial Ethnic Breakdown 

## Black Infant Mortality 

## Children of Incarcerated 

## Foster Youth Population 

## Student Homelessness 

## Homelessness 

## Children with Special Needs 

## Languages of Children and Families
```{r}
source("W:/Project/RDA Team/Region 5 State of the Child/R/language.R")

data
```

# Resources(and barriers to access)

# Economics 

## Rent Burden (Cost Burden Renter)

## Housing Insecurity 

## Cost of Living 

## Food Insecurity 

# Education


## Highschool Graduation

## Teacher & Staff Diversity 

## Chronic Absenteeism 

## Suspensions 

## Explusions 

## Early Childhood Edducation 

## 3rd Grade Math Scores

## 3rd Grade English Scores

# Safety and Education 

## Parks and Safe Spaces

## Police Violence

## Traffic Stops
